```{r}
#| results: "asis"
#| echo: false

source("_common.R")
load("./data/dat0_0118.RData")  # dat0, spec4matrix, yml_spec

library("kableExtra")
library("circlize")

footnote1 <- c(
"DNA_MTH1 = Sample Concentration (cp/µg, VG/ug genomic DNA); 
DNA_MTH2 = Sample Concentration (VG/haploid- Normalized to MfTFRC); 
EVV = Total HumanEVV peptide  (ng/mL); 
LPA = Total Human LPA peptide (ng/mL)")

```

# Overall

The study `r yml_spec$STUDYID$values[1]` involves `r length(unique(dat0$USUBJID))` subjects across `r length(unique(dat0$GROUP))` dose groups, encompassing `r length(unique(dat0$PARAMCD))` analytes or parameters and `r length(unique(dat0$MATRIX))` biological matrices. A summary of the dose groups and the associated number of subjects is presented in @tbl-summary-dose-group.

An overall barplot of the mean expression levels of VGC, mRNA, and proteins across all biological matrices, is shown in @fig-barplot-overall. Additionally, barplots illustrating the mean expression levels in Groups for CNS capsids and DM1 capsids are shown in @fig-barplot-cns-capsids and @fig-barplot-dm1-capsids, respectively.

Tentative observations are summarized below. The more detailed results are presented in the subsequent sections.

**CNS Capsids**

-   The expression levels of VGC and mRNA in CNS-related regions are higher in Groups 3 and 4, which utilize the new JBC capsids developed by JCR Pharmaceuticals, compared to the established control capsid group (Group 2).

-   The capsid used in Group 4, JBC-AAVdQ, is a liver-detargeting variant and is expected to exhibit reduced liver distribution compared to the capsids used in Groups 2 and 3.

-   Distribution in non-target tissues is anticipated to be lower compared to CNS-related regions in Groups 4, 5, and 6.

-   Note that proteins are not detected in some regions or biological matrices due to: 1) the absence of sample collection, or 2) the lack of protein expression in those regions.

**DM1 Capsids**

-   The expression levels of VGC and mRNA in DM1-related regions are higher in Group 6, utilizing the new Gcap-106 capsid, compared to the established control capsid group (Group 5).

-   Distribution in non-target tissues is expected to be lower compared to DM1-related regions in Groups 5 and 6.

-   Note that proteins are not detected in some regions or biological matrices due to: 1) the absence of sample collection, or 2) the lack of protein expression in those regions.

Due to the limited number of subjects in the study, inter-subject variability could not be assessed. Consequently, the above observations are preliminary and require further validation in a larger cohort.

```{r}
#| label: tbl-summary-dose-group
#| tbl-cap: Summary of Dose Groups in the Study
#| tbl-cap-location: top
#| tbl-alignment: "l"
#| 

#install.packages("kableExtra")
library(kableExtra)
library(flextable)
library(knitr)
 
# Summarize data
summary_data <- dat0 %>%
  group_by(GROUP, GROUP_f, DOSE_f) %>%
  dplyr::summarise(N = n_distinct(USUBJID)) %>%
  ungroup()

# Create a flextable with merged cells and zebra striping
ft <- summary_data %>%
  knitr::kable() %>%
  kableExtra::kable_styling(position = "left") 

  # flextable() %>%
  # flextable::merge_v(j = c("GROUP", "GROUP_f", "DOSE_f")) %>%
  # flextable::hline(part = "all", border = officer::fp_border(color = "gray"))

# Print the flextable
ft


 
```

```{r}
#| label: fig-barplot-overall
#| echo: false
#| fig-cap: |
#|   Barplot of Mean Expression Level in VGC, mRNA, LPA and EVV Across All Biological Matrix.
#| fig-cap-location: top
#| fig-width: 8
#| fig-height: 15  
 
 
###############################################
# Run barplot 
###############################################
 
  PARAMCD_filter = c("VGC_MTH1", "RNA_MTH1", "EVV", "LPA" )

  df = dat0   %>% #filter(!GROUP %in% c(1, 2))
  mutate(AVAL = as.numeric(AVAL))  %>% 
  filter(PARAMCD %in%  PARAMCD_filter)    #  "PK"       "VGC_MTH1" "VGC_MTH2" "RNA_MTH1" "RNA_MTH2" "EVV"      "LPA" 
   
  my_barplot(df, facet_by = "PARAMCD") 

  # my_barplot(
  # df, 
  # facet_by = "PARAMCD")  + 
  #  facet_grid(vars(PARAMCD), vars(TISSUECD), scales="free_x")
  
  # theme(
  #   text = element_text(size = 12),  # Adjust text size
  #   axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1),
  #   legend.position = "bottom",
  #   legend.box = "vertical"
  # )


  # my_barplot(
  # df %>% filter(TISSUECD %in% c("BR", "SC", "DRG", "LIV")), 
  # facet_by = "PARAMCD")  + 
  #  facet_grid(vars(PARAMCD), vars(TISSUECD), scales="free_x")
    
# note for all biological matrix
a_matrix_lst <- df %>% #filter(!GROUP %in% c(1, 2))  %>% 
  dplyr::filter(
    PARAMCD %in% PARAMCD_filter    #, ATPT_f %in% ATPT_f_filter
  ) %>%  
  pull(MATRIX)  %>% unique()

abbreviation_of_matrix_lst <- spec4matrix %>%
  dplyr::filter(MATRIX %in% a_matrix_lst) %>% 
  dplyr::arrange(MATRIXCD) %>% 
  tidyr::unite("ABBR", MATRIXCD, MATRIX, sep = " = ") %>% 
  dplyr::pull(ABBR) %>% paste0( collapse="; ")   
 

```

::: {style="color: gray; font-size: 0.8em;"}
Note, `r abbreviation_of_matrix_lst`.

`r footnote1`.
:::

```{r}
#| label: fig-barplot-cns-capsids
#| echo: false
#| fig-cap: |
#|   Barplot of Mean Expression Level in VGC, mRNA, LPA and EVV Across Biological Matrix in CNS-Related Tissue and Group for CNS Capsids (Group 2, 3 and 4) 
#| fig-cap-location: top
#| fig-width: 8
#| fig-height: 15  
 
  GROUP_filter = seq(1, 4, by=1)  # 2, 3, 4   Groups evaluating CNS capsids
  PARAMCD_filter = c("VGC_MTH1", "RNA_MTH1", "EVV", "LPA" )
  TISSUECD_filter = c("BR", "SC")

  df = dat0 %>% 
    filter(GROUP %in% GROUP_filter) %>% 
    filter(PARAMCD %in%  PARAMCD_filter) %>%   #  "PK"       "VGC_MTH1" "VGC_MTH2" "RNA_MTH1" "RNA_MTH2" "EVV"      "LPA" 
    filter(TISSUECD %in%  TISSUECD_filter) %>% 
    mutate(AVAL = as.numeric(AVAL))  
    
library(ggh4x)   
  
  my_barplot(df, facet_by = "PARAMCD")  + 
    facet_grid(vars(PARAMCD), vars(TISSUE), scales="free_x") + 
    labs(fill = "Treatment") + 
    ggh4x::force_panelsizes(cols = c(1, 0.3))  
     
# note for all biological matrix
a_matrix_lst <- df %>% pull(MATRIX)  %>% unique()

abbreviation_of_matrix_lst <- spec4matrix %>%
  dplyr::filter(MATRIX %in% a_matrix_lst) %>% 
  dplyr::arrange(MATRIXCD) %>% 
  tidyr::unite("ABBR", MATRIXCD, MATRIX, sep = " = ") %>% 
  dplyr::pull(ABBR) %>% paste0( collapse="; ")   

```

::: {style="color: gray; font-size: 0.8em;"}
Note, `r abbreviation_of_matrix_lst`.

`r footnote1`.
:::

```{r}
#| label: fig-barplot-dm1-capsids
#| echo: false
#| fig-cap: |
#|   Barplot of Mean Expression Level in VGC, mRNA, LPA and EVV Across Biological Matrix in Musle-Related Tissue and Group for DM1 Capsids (Group 5 and 6) 
#| fig-cap-location: top
#| fig-width: 8
#| fig-height: 15  
 
  GROUP_filter = c(1, 5, 6)  #  Groups evaluating DM1 capsids
  PARAMCD_filter = c("VGC_MTH1", "RNA_MTH1", "EVV", "LPA" )
  TISSUECD_filter = c("MUS")            #  dat0 %>% distinct(TISSUECD, TISSUE)   %>% view()

  df = dat0 %>% 
    filter(GROUP %in% GROUP_filter) %>% 
    filter(PARAMCD %in%  PARAMCD_filter) %>%   #  "PK"       "VGC_MTH1" "VGC_MTH2" "RNA_MTH1" "RNA_MTH2" "EVV"      "LPA" 
    filter(TISSUECD %in%  TISSUECD_filter) %>% 
    mutate(AVAL = as.numeric(AVAL))  
    
  library(ggh4x)   
  
  my_barplot(df, facet_by = "PARAMCD")  + 
    facet_grid(vars(PARAMCD), vars(TISSUE), scales="free_x") + 
    labs(fill = "Treatment")   #+ 
    #ggh4x::force_panelsizes(cols = c(1, 0.3))  
     
  # note for all biological matrix
  a_matrix_lst <- df %>% pull(MATRIX)  %>% unique()
  
  abbreviation_of_matrix_lst <- spec4matrix %>%
    dplyr::filter(MATRIX %in% a_matrix_lst) %>% 
    dplyr::arrange(MATRIXCD) %>% 
    tidyr::unite("ABBR", MATRIXCD, MATRIX, sep = " = ") %>% 
    dplyr::pull(ABBR) %>% paste0( collapse="; ")   

```

::: {style="color: gray; font-size: 0.8em;"}
Note, `r abbreviation_of_matrix_lst`.

`r footnote1`.
:::
